---
title: "Optimisation"
author: "Luke Zappia"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: yes
    theme: yeti
    toc: yes
    toc_float: yes
    df_print: paged
---

Code version: `r system("git log -1 --format=oneline | cut -d' ' -f1", intern = TRUE)`

```{r knitr, include = FALSE}
DOCNAME = "optimisation"
knitr::opts_chunk$set(autodep        = TRUE,
                      cache          = FALSE,
                      cache.path     = paste0("cache/", DOCNAME, "/"),
                      cache.comments = TRUE,
                      echo           = FALSE,
                      error          = FALSE,
                      fig.align      = "center",
                      fig.path       = paste0("figures/", DOCNAME, "/"),
                      fig.width      = 10,
                      fig.height     = 8,
                      message        = FALSE,
                      warning        = FALSE)
```

```{r libaries, cache = FALSE}

```

```{r source, cache = FALSE}

```

Introduction
============

The `aflelo` package contains functions for optimising parameters of the Elo
model using a genetic algorithm. In this document we are going to look at the
results of running this optimisation. There two thing we can try to optimise
for: the error in predicting the margin of games or the percentage of games
where the winner has been correctly predicted. I have run the optimisation three
times, once aiming for the minimum margin error, once aiming for the greatest
win percentage and once with a 50/50 balance between the two. Let's have a look
at the different parameters that have been selected and compare their
performance to the default parameters, based on those used by The Arc.

Session info
============

```{r session-info, cache = FALSE}
devtools::session_info()
```

```{r cleanup-docs, cache = FALSE}
doc.files <- c(list.files(pattern = "pdf"),
               list.files(pattern = "html"),
               list.files(pattern = "docx"))

for (file in doc.files) {
    file.rename(file, file.path("../docs", file))
}
```
